Drug efficacy prediction in cell lines

### 2018-12-20 docs/manuscript.Rnw
independent beta and sigma^2 vb update implemented and tested
created supplementary material with first simulation results

	- IG hyperprior used in finance \cite[]{barndorff-nielsen_normal_1997}, where it is called the NIG prior.
%	\begin{subequations}\label{eq:hyperprior}
%		\begin{align}
%		\sigma_d^{2} & \sim 1/\sigma_d^{3} \\
%		\log \gamma_d^{2} & \sim \mathcal{N}(\theta, \tau^2), \\
%		\bm{\Xi}_d & \sim \mathcal{W}^{-1}_T(\Vm, k),
%		\end{align}
%	\end{subequations}
%- try log-normal distribution for the $\gamma_d$'s. This is used in Example 10.4.1 in \cite{lunn_bugs_2012}, with justification same as ours: allows for modelling the variance as a function of covariates. Also used in examples 6.10 and 9.2 in \cite{abrams_bayesian_2004}.

%	\begin{subequations}\label{eq:hyperprior}
%		\begin{align}
%		\sigma_d^{-2} & \sim \Gamma(0.001, 0.001), \\
%		\gamma_d^{-2} & \sim \Gamma(k, \theta), \\
%		\bm{\Xi}_d & \sim \mathcal{W}^{-1}_T(\Vm, \tau),
%		\end{align}
%	\end{subequations}
%	where $g(j)$ denotes the group index of molecular feature $j$, $\Gamma(k, \theta)$ is a gamma distribution with shape $k$ and rate $\theta$, and $\mathcal{W}_T(\Vm, \tau)$ is a Wishart distribution with $T \times T$-dimensional scale matrix $\Vm$ and degrees of freedom $\tau$. Note that this Wishart prior implicates an expected prior covariance matrix $\tau \cdot \Vm$ for the tissue effects, common to all drugs.

@@

	\subsubsection{Drug-specific variance components}
	The second sub-problem concerns the prior for the drug-specific variance components $\gamma_d^2$. As explained in Section \ref{sec:SEM}, we may have `covariates' $\mathbf{c}_d$ on the drugs available. We conjecture that these covariates may be informative for the omics effect size and therefore model the prior mean of the $\gamma_d^2$ as a function of these covariates: $\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$. Then, the empirical Bayes estimation of $\bm{\alpha}$ effectively boils down to an inverse Gaussian regression of the $\gamma_d^{-2}$ on the $\mathbf{c}_d$ \cite[]{fries_optimal_1986,whitmore_regression_1983}. The problem may be formulated as:
	\begin{align*}\label{eq:optproblem2}
	\lambda^{(l+1)}, \bm{\alpha}^{(l+1)} & = \underset{\lambda,\bm{\alpha}}{\argmax} \E_{Q^{(l)}} [\log \pi (\bgamma^2)] \\
	%	& = \underset{\lambda,\bm{\alpha}}{\argmax} \frac{D}{2} \log \lambda - \frac{\lambda}{2} \sum_{d=1}^D \[ \E(\gamma_d^2)(\mathbf{c} _d \tr \bm{\alpha})^2 - 2 \mathbf{c}_d \tr \bm{\alpha} + \E(\gamma_d^{-2}) \] \\
	& = \underset{\lambda,\bm{\alpha}}{\argmax} \frac{D}{2} \log \lambda - \frac{\lambda}{2} \bm{\alpha} \tr \mathbf{C} \tr \diag [\E_{Q^{(l)}}(\gamma_d^2)] \mathbf{C} \bm{\alpha} + \lambda \bm{\alpha} \tr \mathbf{C} \tr \mathbf{1} - \frac{\lambda}{2} \sum_{d=1}^D \E_{Q^{(l)}}(\gamma_d^{-2}),
	\end{align*}
	where $\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & \mathbf{c}_D \end{bmatrix} \tr$. The solutions to this problem are:
	\begin{align*}
	\bm{\alpha}^{(l+1)} & = \left\{ \mathbf{C} \tr \Vm^{(l)} \mathbf{C} \right\}^{-1} \mathbf{C} \tr \mathbf{1}, \\
	\lambda^{(l+1)} & = \[ \sum_{d=1}^D a_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D,
	\end{align*}
	where 
	$$
	\Vm^{(l)} = \diag \[ \sqrt{\frac{\delta_d^{(l)}(\theta_d^{(l)})^2}{\lambda^{(l)}}} \frac{K_{\frac{p - 1}{2}} \( \sqrt{\lambda^{(l)} \delta_d^{(l)}}/\theta^{(l)}_d \)}{K_{\frac{p + 1}{2}} \( \sqrt{\lambda^{(l)} \delta_d^{(l)}}/\theta^{(l)}_d \)} \].
	$$

@@

%	\begin{align*}
%	\tau^{(l+1)}, \rho^{(l+1)} & = \underset{\tau, \rho}{\argmax} \E_{Q^{(l)}} [\log \pi (\bm{\Xi}_1, \dots, \bm{\Xi}_D)] \\
%	& = \underset{\tau, \rho}{\argmax} w_1^{(l)} \tau - w_2^{(l)} \tau^2 - w_3^{(l)} \tau \rho - D \log \Gamma_T \(\frac{\tau}{2}\) + \frac{DT}{2} \tau \log \tau \\
%	& \,\,\,\,\,\,\,\,\,\,+ \frac{D}{2} \tau \log \[ \tau - (1-T)\rho \] + \frac{D(T-1)}{2} \tau \log (\tau - \rho), \\
%	& \text{subject to } \tau > \max\[\rho - T \rho,\rho, 0\] \text{ and } 1> |\rho| \geq 0,
%	\end{align*}
%	where 
%	\begin{align*}
%	w_1^{(l)} & = \frac{1}{2} \[D \psi_T\(\frac{\tau^{(l)} + 1}{2}\) + \sum_{d=1}^D \log |\bm{\Psi}_d^{(l)}| \], \\
%	w_2^{(l)} & = \frac{\tau^{(l)} + 1}{2} \sum_{d=1}^D \trace \(\bm{\Psi}_d^{(l)}\) > 0,\\
%	w_3^{(l)} & = \frac{\tau^{(l)} + 1}{2} \sum_{d=1}^D \trace \[(\mathbf{1}_{T \times T} - \I_T) \cdot \bm{\Psi}_d^{(l)}\],
%	\end{align*}
%	and $\Gamma_T(x)$ and $\psi_T(x)$ are the $T$-variate gamma and digamma functions, respectively. $\log \Gamma_T \(\frac{\tau}{2}\)$ is convex on the intervals $(0,1), \dots, (T-1, \infty)$

%	- we model the prior expectation of the covariance instead of the precision. Precisions are conditional on the other variables, which doesn't make sense in the tissue context.

%	The second sub problem estimates the molecular feature effect hyper prior:
%	\begin{align}\label{eq:optproblem2}
%	k^{(l+1)},\theta^{(l+1)} & = \underset{k,\theta}{\argmax} \E_{Q^{(l)}} [\log \pi (\bgamma^2)] \nonumber\\
%	& = \underset{k,\theta}{\argmax} D k \log \theta - D \log \Gamma (k) + e_1^{(l)} k - e_2^{(l)} \theta \\
%	& \text{subject to } k ,\theta > 0,
%	\end{align}
%	where
%	\begin{align*}
%	e_1^{(l)} & = D\psi\(\frac{p}{2} + k^{(l)}\) - \sum_{d=1}^D \log a_d^{(l)}, \\
%	e_2^{(l)} & = \(\frac{p}{2} + k^{(l)} \) \sum_{d=1}^D (a_d^{(l)})^{-1},
%	\end{align*}
%	and $\Gamma(x)$ and $\psi(x)$ are the gamma and digamma functions, respectively. This is again a convex problem (see Appendix \ref{app:convexity}), but more cumbersome to solve numerically due to the log-gamma function. By approximating the digamma function with $\phi(x) \approx \log x - 1/(2x)$, we get the following closed-form solutions:
%	\begin{align*}
%	k^{(l+1)} & = \frac{1}{2} \( \log e_2^{(l)} - D^{-1} e_1^{(l)} - \log D \)^{-1}, \\
%	\theta^{(l+1)} & = D k^{(l+1)} (e_2^{(l)})^{-1}.
%	\end{align*}

@@

	\titleformat{\section}{\normalfont\Large\bfseries}{\appendixname~\thesection.}{1em}{}
	\begin{appendix}
		\section{Variational Bayes derivations}\label{app:derivations}
		In the following all expectations are with respect to the variational posterior $Q$.
		\begin{align*}
		\log q_{\B} (\B) & \propto \E_{\mathbf{U}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\bgamma^2, \bsigma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
		& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\mathbf{u}_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E_{\gamma^2_d, \sigma^2_d} \( \frac{\beta_{jd}^2}{\gamma_d^2 \sigma_d^2} \), \\
		q_{\B} (\B) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_p (\bmu_d, \bSigma_d), \\
		& \text{ with } \bSigma_d = \E(\sigma_d^{-2})^{-1} [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1}, \\
		& \text{ and } \bmu_d = [\X \tr \X + \E(\gamma_d^{-2}) \cdot \diag (\phi_{g(j)}^{-2}) ]^{-1} \X \tr [\y_d - \Z \cdot \E(\mathbf{u}_d)].
		\end{align*}

		\begin{align*}
		\log q_{\U}(\U) & \propto \E_{\mathbf{B}, \bsigma^2} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2} [\log \pi (\U | \mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D, \bsigma^2)] \\
		& \propto - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E_{\bbeta_d, \sigma_d^2} \[ \frac{(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2}{\sigma_d^2} \] - \frac{1}{2} \sum_{d=1}^D \E_{\bm{\Xi}_d, \sigma_d^2}\( \frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \), \\
		q_{\U}(\U) & \overset{D}{=} \prod_{d=1}^{D} \mathcal{N}_T (\mathbf{m}_d, \Sm_d), \\
		& \text{ with } \Sm_d = \E(\sigma_d^{-2})^{-1} [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1}, \\
		& \text{ and } \mathbf{m}_d = [\Z \tr \Z + \E(\bm{\Xi}_d^{-1}) ]^{-1} \Z \tr [\y_d - \X \cdot \E(\bbeta_d)].
		\end{align*}

		\begin{align*}
		\log q_{\bgamma^{2}}(\bgamma^{2}) & \propto \E_{\B, \bsigma^2} [\log \pi (\B | \bgamma^{2}, \sigma^2)] + \log \pi (\bgamma^{2}) \\
		& \propto -\frac{p}{2} \sum_{d=1}^D \log \gamma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_d^{-2} - \frac{3}{2} \sum_{d=1}^D \log \gamma_d^{2} \\
		& \,\,\,\,\,\,\,\,\,\, - \frac{\lambda}{2 \theta^2} \sum_{d=1}^D (\gamma_d^2 - \theta)^2 \gamma_d^{-2} \\
		& \propto \sum_{d=1}^D \left\{ \( -\frac{p + 1}{2} - 1\) \log \gamma_d^{2} - \frac{\lambda}{2\theta^2} \gamma_d^2 - \frac{1}{2}\[ \lambda + \E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^{2}}\] \gamma_d^{-2} \right\}, \\
		q_{\bgamma^{2}}(\bgamma^{2}) & \overset{D}{=} \prod_{d=1}^D \mathcal{GIG} \(-\frac{p+1}{2}, \frac{\lambda}{\theta_d^2}, \delta_{d} \), \\
		& \text{with } \delta_{d} =\E(\sigma_d^{-2}) \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} + \lambda.
		\end{align*}

%		\begin{align*}
%		\log q_{\bgamma^{-2}}(\bgamma^{-2}) & \propto \E_{\B, \bsigma^2} [\log \pi (\B | \bgamma^{-2}, \sigma^2)] + \log \pi (\bgamma^{-2}) \\
%		& \propto \frac{p}{2} \sum_{d=1}^D \log \gamma_d^{-2} - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E \( \frac{\beta_{jd}^2}{\sigma^{2}_d}\) \gamma_d^{-2} + (k - 1) \sum_{d=1}^D \log \gamma_d^{-2} - \theta \sum_{d=1}^D \gamma_d^{-2} \\
%		& = \sum_{d=1}^D \left\{ \( \frac{p}{2} + k - 1\) \log \gamma_d^{-2} - \[ \frac{\E(\sigma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^{2}} + \theta \] \gamma_d^{-2} \right\}.
%		\end{align*}

%	\begin{align*}
%	q_{\bgamma^{-2}}(\bgamma^{-2}) & \overset{D}{=} \prod_{d=1}^D \Gamma \(\frac{p}{2} + k, a_{d} \), \\
%	& \text{ with } a_{d} =\frac{\E(\sigma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} + \theta.
%	\end{align*}

		\begin{align*}
		\log q_{\bsigma^{2}}(\bsigma^{2}) & \propto \E_{\B,\mathbf{U}} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\B, \bgamma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
		& \,\,\,\,\,\,\,\,\,\, + \E_{\mathbf{U}, \bm{\Xi}_1, \dots, \bm{\Xi}_D} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bsigma^{2}) \\
		& \propto -\frac{n}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \sigma_d^{2} - \frac{p}{2} \sum_{d=1}^D \log \sigma_d^{-2} \\
		& \,\,\,\,\,\,\,\,\,\, - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \E( \gamma_d^{-2}) \frac{\E ( \beta_{jd}^2)}{\phi_{g(j)}^{2}} \sigma_d^{-2} - \frac{T}{2} \sum_{d=1}^D \log \sigma_d^{2} - \frac{1}{2} \sum_{d=1}^D \E(\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}) \sigma_d^{-2} \\
		& \,\,\,\,\,\,\,\,\,\, - \frac{3}{2} \sum_{d=1}^D \log \sigma^2_d \\
		& = -\sum_{d=1}^{D} \(\frac{n + p + T + 3}{2} \) \log \sigma_d^{2} - \sum_{d=1}^{D} \frac{1}{2}\Bigg\{ \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \\
		& \,\,\,\,\,\,\,\,\,\, + \E(\gamma_d^{-2})\sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^2} + \E(\mathbf{u}_{d} \tr \bm{\Xi}_d^{-1} \mathbf{u}_d) \Bigg\} \sigma_d^{-2}, \\
		q_{\bsigma^{2}}(\bsigma^{2}) & \overset{D}{=} \prod_{d=1}^D \Gamma^{-1} \(\frac{n + p + T + 1}{2}, \zeta_{d} \), \\
		& \text{ with } \zeta_{d} = \frac{\y_d \tr \y_d}{2} - \y_d \tr [\X \E(\bbeta_d) + \Z \E(\mathbf{u}_d)] + \E(\bbeta_d \tr) \X \tr \Z \E (\mathbf{u}_d) \\
		& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\trace [\X \tr \X \V(\bbeta_d)]}{2} + \frac{\E(\bbeta_d \tr) \X \tr \X \E(\bbeta_d)}{2} + \frac{\trace [\Z \tr \Z \V(\mathbf{u}_d)]}{2} \\
		& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\E(\mathbf{u}_d \tr) \Z \tr \Z \E(\mathbf{u}_d)}{2} + \frac{\E(\gamma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} \\
		& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,+ \frac{\E(\mathbf{u}_d \tr) \E(\bm{\Xi}_d^{-1}) \E(\mathbf{u}_d)}{2} + \frac{\trace [\E(\bm{\Xi}_d^{-1}) \V(\mathbf{u}_d)]}{2},
		\end{align*}

%		\begin{align*}
%		\log q_{\bsigma^{-2}}(\bsigma^{-2}) & \propto \E_{\B,\mathbf{U}} [\log \mathcal{L}(\Y | \B, \U, \bsigma^2 )] + \E_{\B, \bgamma^2} [\log \pi (\B | \bgamma^2, \bsigma^2)] \\
%		& \,\,\,\,\,\,\,\,\,\, + \E_{\mathbf{U}, \bm{\Xi}_1, \dots, \bm{\Xi}_D} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bsigma^{-2}) \\
%		& \propto \frac{n}{2} \sum_{d=1}^D \log \sigma_u^{-2} - \frac{1}{2} \sum_{d=1}^D \sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2] \sigma_d^{-2} + \frac{p}{2} \sum_{d=1}^D \log \sigma_d^{-2} \\
%		& \,\,\,\,\,\,\,\,\,\, - \frac{1}{2} \sum_{d=1}^D \sum_{j=1}^p \phi_{g(j)}^{-2} \E \( \frac{\beta_{jd}^2}{\gamma_d^{2}} \) \sigma_d^{-2} + \frac{T}{2} \sum_{d=1}^D \log \sigma_d^{-2} - \frac{1}{2} \sum_{d=1}^D \E(\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}) \sigma_d^{-2} \\
%		& \,\,\,\,\,\,\,\,\,\, - 0.999 \sum_{d=1}^D \log \sigma_d^{-2} - 0.001 \sum_{d=1}^D \sigma_d^{-2} \\
%		& = \sum_{d=1}^{D} \(\frac{n + p + T}{2} - 0.999 \) \log \sigma_d^{-2} - \sum_{d=1}^{D} \Bigg\{ \frac{\sum_{i=1}^n \E [(y_{id} - \x_i \tr \bbeta_d - \mathbf{z}_i \tr \mathbf{u}_d)^2]}{2} \\
%		& \,\,\,\,\,\,\,\,\,\, + \frac{\E(\gamma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd}^2)}{\phi_{g(j)}^2} + \frac{\E(\mathbf{u}_{d} \tr \bm{\Xi}_d^{-1} \mathbf{u}_d)}{2}  + 0.001 \Bigg\} \sigma_d^{-2}.
%		\end{align*}

%	\begin{align*}
%	q_{\bsigma^{-2}}(\bsigma^{-2}) & \overset{D}{=} \prod_{d=1}^D \Gamma \(\frac{n + p + T}{2} + 0.001, b_{d} \), \\
%	& \text{ with } b_{d} = \frac{\y_d \tr \y_d}{2} - \y_d \tr [\X \E(\bbeta_d) + \Z \E(\mathbf{u}_d)] + \E(\bbeta_d \tr) \X \tr \Z \E (\mathbf{u}_d) \\
%	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\trace [\X \tr \X \V(\bbeta_d)]}{2} + \frac{\E(\bbeta_d \tr) \X \tr \X \E(\bbeta_d)}{2} + \frac{\trace [\Z \tr \Z \V(\mathbf{u}_d)]}{2} \\
%	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + \frac{\E(\mathbf{u}_d \tr) \Z \tr \Z \E(\mathbf{u}_d)}{2} + \frac{\E(\gamma_d^{-2})}{2} \sum_{j=1}^p \frac{\E(\beta_{jd})^2 + \V(\beta_{jd})}{\phi_{g(j)}^2} \\
%	& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,+ \frac{\E(\mathbf{u}_d \tr) \E(\bm{\Xi}_d^{-1}) \E(\mathbf{u}_d)}{2} + \frac{\trace [\E(\bm{\Xi}_d^{-1}) \V(\mathbf{u}_d)]}{2} + 0.001.
%	\end{align*}

		\begin{align*}
		\log q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \propto \E_{\mathbf{U}, \bm{\sigma}^2} [\log \pi (\mathbf{U} | \bm{\Xi}_1, \dots, \bm{\Xi}_D, \bsigma^2)] + \log \pi (\bm{\Xi}_1, \dots, \bm{\Xi}_D) \\
		& \propto -\frac{1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \E \(\frac{\mathbf{u}_d \tr \bm{\Xi}_d^{-1} \mathbf{u}_d}{\sigma_d^2} \) \\
		& \,\,\,\,\,\,\,\,\,\, - \frac{k + T + 1}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace (\bm{\Omega} \bm{\Xi}_d^{-1}) \\
		& = -\frac{\tau + T + 2}{2} \sum_{d=1}^D \log |\bm{\Xi}_d| - \frac{1}{2} \sum_{d=1}^D \trace \left\{ \[\E(\sigma_d^{-2}) \E( \mathbf{u}_d \mathbf{u}_d \tr) + \bm{\Omega}\] \bm{\Xi}_d^{-1} \right\}, \\
		q_{\bm{\Xi}_1, \dots, \bm{\Xi}_D}(\bm{\Xi}_1, \dots, \bm{\Xi}_D) & \overset{D}{\propto} \prod_{d=1}^D \mathcal{W}_T^{-1} (\bm{\Psi}_d, \nu + 1),  \\ 
		& \text{ with } \bm{\Psi}_d =\E(\sigma^{-2}_d) \[\V(\mathbf{u}_d) + \E(\mathbf{u}_d) \E( \mathbf{u}_d \tr) \] + \bm{\Omega}.
		\end{align*}

%		\section{Convexity}\label{app:convexity}
%		For (\ref{eq:optproblem2}) to be a convex problem, we need the Hessian $\mathbf{H}$ of the objective function to be positive definite, i.e., $\forall \mathbf{z} \in \mathbb{R}^2: \mathbf{z} \tr \mathbf{H} \mathbf{z} > 0$. We have
%		\begin{align*}
%		\mathbf{z} \tr \mathbf{H} \mathbf{z} = \frac{D}{\theta^2 \psi^{(1)}(k)} \( z_1 \theta \psi^{(1)}(k) - z_2 \)^2 + z_2^2 \( k \psi^{(1)}(k)  - 1 \),
%		\end{align*}
%		with $\psi^{(1)}(k)$ the trigamma function. Since the trigamma function is positive for all $k > 0$, and we have $\theta, k > 0$, we need:
%		$$
%		k \psi^{(1)}(k)> 1.
%		$$
%		Using Lemma 1 in \cite{guo_refinements_2013} we can show that for $k > 0$:
%		$$
%		k \psi^{(1)}(k) > 1 + \frac{1}{2k},
%		$$
%		such that the objective function is convex.

		\section{Prior tissue covariance constraints}\label{app:priortissuecovariance}
%		The empirical Bayes objective function of the third sub-optimisation problem in \ref{sec:empiricalbayes} is:
%		\begin{align*}
%		\E_{Q^{(l)}}[\log \pi (\mathbf{\Xi}_1, \dots, \mathbf{\Xi}_D)] & \propto \frac{D}{2} \nu \log |\bm{\Omega}| - \frac{D T \log 2}{2} \nu - D \log \Gamma_T \(\frac{\nu}{2}\) + \frac{\nu}{2}\sum_{d=1}^D \E_{Q^{(l)}} (\log | \mathbf{\Xi}_d^{-1}|) \\
%		& \,\,\,\,\,\,\,\,\,\, - \frac{1}{2} \sum_{d=1}^D \trace [ \bm{\Omega} \cdot \E_{Q^{(l)}}(\mathbf{\Xi}_d^{-1})] \\
%		& = \frac{D}{2} \nu \log |\bm{\Omega}| - D \log \Gamma_T\(\frac{\nu}{2}\) - \frac{\nu^{(l)} + 1}{2} \sum_{d=1}^D \trace [ \bm{\Omega} \cdot ( \mathbf{\Psi}_d^{(l)})^{-1}] \\
%		& \,\,\,\,\,\,\,\,\,\, + \frac{\nu}{2} \sum_{d=1}^D \[ \psi_T\(\frac{\nu^{(l)} + 1}{2}\) - \log |\mathbf{\Psi}_d^{(l)}| \],
%		\end{align*}
%		which is maximised under constraints $\nu > T + 1$ and $\bm{\Omega}$ positive definite (PD), to ensure a proper inverse Wishart distribution.

		We parametrise the prior covariance as a matrix with $\tau$ on the diagonal and $\rho$ on its off-diagonals. To that end we set $\nu = \tau + T + 1$ and the $T \times T$-dimensional matrix
		$$
		\bm{\Omega} = \tau \begin{bmatrix}
		\tau &  & \rho \\
		& \ddots &  \\
		\rho & & \tau
		\end{bmatrix},
		$$
		such that $\E_{\pi(\mathbf{\Xi}_d)} (\mathbf{\Xi}_d) = (\nu - T - 1)^{-1} \bm{\Omega}$ is of the desired form. To ensure a non-degenerate prior distribution we require $\tau > 0$ and $\bm{\Omega}$ PD. We use $\tau > 0$ to note that a PD $\bm{\Omega}$ implies positive roots $\lambda$ of the characteristic polynomial:
		\begin{align*}
		\left| \begin{bmatrix}
		\tau &  & \rho \\
		& \ddots &  \\
		\rho & & \tau
		\end{bmatrix} -\lambda \I_T \right| = (\tau - \lambda - \rho)^{T-1}[\tau - \lambda - (1-T)\rho] = 0.
		\end{align*}
		For $T > 1$, the solutions are:
		$$
		(\lambda = \tau - \rho) \lor (\lambda = T \rho - \rho + \tau),
		$$
		which are positive if the following condition holds: $\tau > \max(\rho,\rho - T \rho)$. We combine this with $\tau > 0$, to arrive at the full proper prior constraint: $\tau > \max(0,\rho,\rho - T \rho)$.

%		The resulting empirical Bayes step is now:
%		\begin{align*}
%		\tau^{(l + 1)}, \rho^{(l+1)} & = \underset{\tau, \rho}{\argmax} \frac{D (T^2 + T)}{2} \log \tau + \frac{DT}{2} \tau \log \tau + \frac{D(T+1)}{2} \log \[ \tau - (T - 1) \rho\] \\
%		& \,\,\,\,\,\,\,\,\,\, + \frac{D}{2} \tau \log \[ \tau - (T - 1) \rho\] + \frac{D(T^2 - 1)}{2} \log(\tau - \rho) +\frac{D(T - 1)}{2} \tau  \log(\tau - \rho) \\
%		& \,\,\,\,\,\,\,\,\,\, - D \sum_{t=1}^T \log \Gamma \( \frac{\tau + t + 1}{2} \) - w_1^{(l)} \tau^2 + w_2^{(l)} \tau \rho + w_3^{(l)} \tau, \\
%		& \text{subject to } \tau > \max\[\rho - T \rho,\rho, 0\],
%		\end{align*}
%		with
%		\begin{align*}
%		w_1^{(l)} & =  \frac{\tau^{(l)} + T + 2}{2} \sum_{d=1}^D \trace \[ (\bm{\Psi}_d^{(l)})^{-1}\] > 0,\\
%		w_2^{(l)} & = \frac{\tau^{(l)} + T + 2}{2} \sum_{d=1}^D \left\{\trace \[ (\bm{\Psi}_d^{(l)})^{-1}\] - \trace \[\mathbf{1}_{T \times T} \cdot (\bm{\Psi}_d^{(l)})^{-1}\] \right\} < w_1^{(l)},\\
%		w_3^{(l)} & = \frac{1}{2} \[D \sum_{t=1}^T \psi\(\frac{\tau^{(l)} + t + 2}{2}\) - \sum_{d=1}^D \log |\bm{\Psi}_d^{(l)}| \].
%		\end{align*}

>>>>

	- IG hyperprior used in finance \cite[]{barndorff-nielsen_normal_1997}, where it is called the NIG prior.

@@

	\subsubsection{Drug-specific variance components}
	The second sub-problem concerns the prior for the drug-specific variance components $\gamma_d^2$. As explained in Section \ref{sec:SEM}, we may have `covariates' $\mathbf{c}_d$ on the drugs available. We conjecture that these covariates may be informative for the omics effect size and therefore model the prior mean of the $\gamma_d^2$ as a function of these covariates: $\theta_d = (\mathbf{c}_d \tr \bm{\alpha})^{-1}$. Then, the empirical Bayes estimation of $\bm{\alpha}$ effectively boils down to an inverse Gaussian regression of the $\gamma_d^{-2}$ on the $\mathbf{c}_d$ \cite[]{fries_optimal_1986,whitmore_regression_1983}. The problem may be formulated as:
	\begin{align*}\label{eq:optproblem2}
	\lambda^{(l+1)}, \bm{\alpha}^{(l+1)} & = \underset{\lambda,\bm{\alpha}}{\argmax} \E_{Q^{(l)}} [\log \pi (\bgamma^2)] \\
	& = \underset{\lambda,\bm{\alpha}}{\argmax} \frac{D}{2} \log \lambda - \frac{\lambda}{2} \bm{\alpha} \tr \mathbf{C} \tr \diag [\E_{Q^{(l)}}(\gamma_d^2)] \mathbf{C} \bm{\alpha} + \lambda \bm{\alpha} \tr \mathbf{C} \tr \mathbf{1} - \frac{\lambda}{2} \sum_{d=1}^D \E_{Q^{(l)}}(\gamma_d^{-2}),
	\end{align*}
	where $\mathbf{C} = \begin{bmatrix} \mathbf{c}_1 & \dots & \mathbf{c}_D \end{bmatrix} \tr$. The solutions to this problem are:
	\begin{align*}
	\bm{\alpha}^{(l+1)} & = \left\{ \mathbf{C} \tr \Vm^{(l)} \mathbf{C} \right\}^{-1} \mathbf{C} \tr \mathbf{1}, \\
	\lambda^{(l+1)} & = \[ \sum_{d=1}^D a_d^{(l)} - (\bm{\alpha}^{(l+1)}) \tr \mathbf{C} \tr \mathbf{1} \]^{-1} D,
	\end{align*}
	where 
	$$
	\Vm^{(l)} = \diag \[ \sqrt{\frac{\delta_d^{(l)}(\theta_d^{(l)})^2}{\lambda^{(l)}}} \frac{K_{\frac{p - 1}{2}} \( \sqrt{\lambda^{(l)} \delta_d^{(l)}}/\theta^{(l)}_d \)}{K_{\frac{p + 1}{2}} \( \sqrt{\lambda^{(l)} \delta_d^{(l)}}/\theta^{(l)}_d \)} \].
	$$

@@

@@

	\titleformat{\section}{\normalfont\Large\bfseries}{\appendixname~\thesection.}{1em}{}
	\begin{appendix}
		\section{Prior tissue covariance constraints}\label{app:priortissuecovariance}
		We parametrise the prior covariance as a matrix with $\tau$ on the diagonal and $\rho$ on its off-diagonals. To that end we set $\nu = \tau + T + 1$ and the $T \times T$-dimensional matrix
		$$
		\bm{\Omega} = \tau \begin{bmatrix}
		\tau &  & \rho \\
		& \ddots &  \\
		\rho & & \tau
		\end{bmatrix},
		$$
		such that $\E_{\pi(\mathbf{\Xi}_d)} (\mathbf{\Xi}_d) = (\nu - T - 1)^{-1} \bm{\Omega}$ is of the desired form. To ensure a non-degenerate prior distribution we require $\tau > 0$ and $\bm{\Omega}$ PD. We use $\tau > 0$ to note that a PD $\bm{\Omega}$ implies positive roots $\lambda$ of the characteristic polynomial:
		\begin{align*}
		\left| \begin{bmatrix}
		\tau &  & \rho \\
		& \ddots &  \\
		\rho & & \tau
		\end{bmatrix} -\lambda \I_T \right| = (\tau - \lambda - \rho)^{T-1}[\tau - \lambda - (1-T)\rho] = 0.
		\end{align*}
		For $T > 1$, the solutions are:
		$$
		(\lambda = \tau - \rho) \lor (\lambda = T \rho - \rho + \tau),
		$$
		which are positive if the following condition holds: $\tau > \max(\rho,\rho - T \rho)$. We combine this with $\tau > 0$, to arrive at the full proper prior constraint: $\tau > \max(0,\rho,\rho - T \rho)$.
